### 2018-05-11 docs/google-research-awards-latam/bug_table.tex
pass on results

\begin{table}[]
    \centering
    \caption{Bugs found and reported by \Fix{JSFuzz} team}
    \label{tab:bugs}
    \begin{tabular}{|c|c|c|c|}
    \hline
    Issue    & Engine                                                          & Status                                                                            & \multicolumn{1}{c|}{Url}                                                                                                                      \\ \hline
    Bug \#1  & Chakra                                                          & Fixed                                                                             & \href{https://github.com/Microsoft/ChakraCore/issues/4978}{\#4978}                                                                                           \\ \hline
    Bug \#2  & Chakra                                                          & Rejected                                                                          & \href{https://github.com/Microsoft/ChakraCore/issues/4979}{\#4979}                                                                              \\ \hline
    Bug \#3  & JavascriptCore                                                  & New                                                                               & \href{https://bugs.webkit.org/show\_bug.cgi?id=184629}{\#184629}                                                                                               \\ \hline
    Bug \#4  & JavascriptCore                                                  & New                                                                               & \href{https://bugs.webkit.org/show\_bug.cgi?id=184749}{\#184749}                                                                                               \\ \hline
    Bug \#5  & Chakra                                                          & Confirmed                                                                         & \href{https://github.com/Microsoft/ChakraCore/issues/5033}{\#5033}                                                                                           \\ \hline
    Bug \#6  & Chakra                                                          & Fixed                                                                             & \href{https://github.com/Microsoft/ChakraCore/issues/5038}{\#5038}                                                                                           \\ \hline
    Bug \#7  & Chakra                                                          & Confirmed                                                                         & \href{https://github.com/Microsoft/ChakraCore/issues/5065}{\#5065}                                                                                           \\ \hline
    Bug \#8  & \begin{tabular}[c]{@{}c@{}}Chakra\\ JavascriptCore\end{tabular} & \begin{tabular}[c]{@{}c@{}}Confirmed (Chakra)\\ New (JavascriptCore)\end{tabular} & \begin{tabular}[c]{@{}l@{}}\href{https://github.com/Microsoft/ChakraCore/issues/5067}{\#5067}\\ \href{https://bugs.webkit.org/show\_bug.cgi?id=185130}{\#185130}\end{tabular} \\ \hline
    Bug \#9  & JavascriptCore                                                  & New                                                                               & \href{https://bugs.webkit.org/show\_bug.cgi?id=185127}{\#185127}                                                                                               \\ \hline
    Bug \#10 & \begin{tabular}[c]{@{}c@{}}Chakra\\ JavascriptCore\end{tabular} & \begin{tabular}[c]{@{}c@{}}Confirmed (Chakra)\\ New (JavascriptCore)\end{tabular} & \begin{tabular}[c]{@{}l@{}}\href{https://github.com/Microsoft/ChakraCore/issues/5076}{\#5076}\\ \href{https://bugs.webkit.org/show\_bug.cgi?id=185156}{\#185156}\end{tabular} \\ \hline
    Bug \#11 & JavascriptCore                                                  & New                                                                               & \href{https://bugs.webkit.org/show\_bug.cgi?id=185197}{\#185197}                                                                                               \\ \hline
    Bug \#12 & JavascriptCore                                                  & New                                                                               & \href{https://bugs.webkit.org/show\_bug.cgi?id=185208}{\#185208}                                                                                               \\ \hline
    Bug \#13 & Chakra                                                          & New                                                                               & \href{https://github.com/Microsoft/ChakraCore/issues/5128}{\#5128}                                                                      \\ \hline
    \end{tabular}
    \end{table}