### 2018-05-11 docs/google-research-awards-latam/bugs.tex

\lstdefinelanguage{JavaScript}{
  keywords={break, case, catch, continue, debugger, default, delete, do, else, finally, for, function, if, in, instanceof, new, return, switch, this, throw, try, typeof, var, void, while, with},
  morecomment=[l]{//},
  morecomment=[s]{/*}{*/},
  morestring=[b]',
  morestring=[b]",
  sensitive=true
}

\begin{figure}[htbp]
    \centering
    \begin{subfigure}[b]{0.4}
        \centering
        \scriptsize
        \lstset{
            language=JavaScript,
            keywords={function, return},
            escapeinside={@}{@},
            numbers=left,xleftmargin=1em,frame=single,framexleftmargin=0.5em,
            basicstyle=\ttfamily\scriptsize, boxpos=c,
            numberstyle=\tiny,
            linewidth=6cm
        }
\begin{lstlisting} 
var a = {
    valueOf: function () {
-    return ""
+    return "\x00"
    }
}
assert(+a === 0)
\end{lstlisting}
        \normalsize
        \caption{
            \label{fig:bug-chakra}Fuzzed file shows an unexpected behaviour on ChakraCore
        }        
    \end{subfigure}
    \begin{subfigure}[b]{0.4}
\begin{lstlisting}[basicstyle=\ttfamily]
Priority: HIGH
Pattern:
-----JavaScriptCore
Error: Test failed
-----Chakra
-----SpiderMonkey
Error: Test failed
-----v8
Error: Test failed
\end{lstlisting}
\caption{\label{fig:pattern}\Fix{JSFuzz} output log}
\end{subfigure}
\end{figure}