### 2019-01-28 paper/table_differential.tex
finding some room for new contributions

\begin{table}[t]
  \small
  \vspace{-3ex}
  \centering
  \caption{List of bug reports issued as result of cross-engine
    differential testing.}
  \label{tab:bugs}
  \setlength{\tabcolsep}{2pt}
  \renewcommand{\arraystretch}{0.95}
  \begin{tabular}{rcccccc}
    \toprule
    Issue\# \Comment{ & Date} & Fuzzer & Engine  & Status \Comment{&
    \multicolumn{1}{c}{Url}}  & Severity & Priority & Suite \\
    \midrule
    1  \Comment{& 4/12} & radamsa & \chakra{}   & \textbf{Fixed}  \Comment{&
    \anonym{\href{https://github.com/Microsoft/\chakra{}Core/issues/4978}{\#4978}}}
    & 2 & \textbf{\lo} & \jsc{} \\ 
    2  \Comment{& 4/12} & radamsa & \chakra{}   & WontFix  \Comment{&
    \anonym{\href{https://github.com/Microsoft/\chakra{}Core/issues/4979}{\#4979}}}
    & - & \hi{} & \jsc{} \\
    3  \Comment{& 4/14} & radamsa & \jsc{}  & New \Comment{&
    \anonym{\href{https://bugs.webkit.org/show\_bug.cgi?id=184629}{\#184629}}
    } & -  & \hi{} & \jsc{}    \\
    4  \Comment{& 4/25} & radamsa & \chakra{}  & \textbf{Fixed}     \Comment{&
    \anonym{\href{https://github.com/Microsoft/\chakra{}Core/issues/5038}{\#5038}}}
    & 2 & \hi{} & \jerry{}   \\
    5  \Comment{& 4/29} & radamsa & \jsc{}  & \textbf{Fixed}  \Comment{&
    \anonym{\href{https://bugs.webkit.org/show\_bug.cgi?id=185127}{\#185127}}
    } & 2  & \hi{}  & \jerry{}\\

    6 \Comment{& 4/30}  & radamsa & \chakra{} & \textbf{Confirmed} \Comment{&
    \anonym{\href{https://github.com/Microsoft/\chakra{}Core/issues/5076}{\#5076}}}
    & 2 & \hi{} & TinyJS\\

    7                    \Comment{& 4/30}                       &  radamsa   &
    \jsc{} & New \Comment{&
    \anonym{\href{https://bugs.webkit.org/show\_bug.cgi?id=185156}{\#185156}}}
    & - & \hi{} & TinyJS \\

    8 \Comment{& 5/02} & radamsa & \jsc{}  & \textbf{Fixed} \Comment{&
    \anonym{\href{https://bugs.webkit.org/show\_bug.cgi?id=185197}{\#185197}}}
    & 2 & \textbf{\lo} & \smonkey{} \\
    9 \Comment{& 5/10} & radamsa & \chakra{} & \textbf{Confirmed} \Comment{&
    \anonym{\href{https://github.com/Microsoft/\chakra{}Core/issues/5128}{\#5128}}}
    & 3 & \hi{} & \jerry{} \\
    10 \Comment{& 5/17} & radamsa & \chakra{} & \textbf{Fixed} \Comment{&
    \anonym{\href{https://github.com/Microsoft/\chakra{}Core/issues/5182}{\#5182}}}
    & 2 & \hi{} & \veight{}\\
    11 \Comment{& 5/24} & radamsa & \jsc{} & \textbf{Fixed}  \Comment{&
    \anonym{\href{https://bugs.webkit.org/show\_bug.cgi?id=185943}{\#185943}}}
    & 2 & \hi{} & \jsc{}\\
    12 \Comment{& 6/26} & radamsa & \jsc{} & \textbf{Fixed}  \Comment{&
    \anonym{\href{https://bugs.webkit.org/show_bug.cgi?id=187042}{\#187042}}}
    & 2 & \hi{} & \jerry{}\\
    13 \Comment{& 7/10} & quickfuzz & \jsc{} & \textbf{Fixed}  \Comment{&
    \anonym{\href{https://bugs.webkit.org/show_bug.cgi?id=187520}{\#187520}}}
    & 2 & \hi{} & \jerry{}\\
    14 \Comment{& 7/10} & quickfuzz & \chakra{} & \textbf{Confirmed}  \Comment{&
    \anonym{\href{https://github.com/Microsoft/\chakra{}Core/issues/5443}{\#5443}}}
    & 2 & \hi{} & \jerry{}\\
    %    \multicolumn{9}{c}{\emph{Non-Exploratory Phase}} \\

    15 \Comment{& 8/21}  & radamsa & \chakra{} & \textbf{Confirmed} \Comment{&
    \anonym{\href{https://github.com/Microsoft/\chakra{}Core/issues/5617}{\#5617}}}
    & - & \hi{} & Test262\\

    16 \Comment{& 8/21} & radamsa &
    \veight{} & \textbf{Fixed} \Comment{&
    \anonym{\href{https://bugs.chromium.org/p/v8/issues/detail?id=8078}{\#8078}}}
    & 2 & \hi{} & Test262 \\

    17 \Comment{& 8/23} & quickfuzz & \jsc{} & New  \Comment{&
    \anonym{\href{https://bugs.webkit.org/show_bug.cgi?id=188899}{\#188899}}}
    & - & \hi{} & Test262\\
    18 \Comment{& 8/23} & quickfuzz & \veight{} & \textbf{Confirmed}  \Comment{&
    \anonym{\href{https://bugs.chromium.org/p/v8/issues/detail?id=8088}{\#8088}}}
    & 2 & \hi{} & Test262\\

    19 \Comment{& 8/24}  & quickfuzz & \chakra{} & New \Comment{&
    \anonym{\href{https://github.com/Microsoft/\chakra{}Core/issues/5630}{\#5630}}}
    & - & \hi{} & Test262\\

    20 \Comment{& 8/24} &  quickfuzz & 
    \jsc{} & New \Comment{&
    \anonym{\href{https://bugs.webkit.org/show_bug.cgi?id=188920}{\#188920}}}
    & - & \hi{} & Test262 \\

    21 \Comment{& 8/24} & quickfuzz & \jsc{}  & New \Comment{&
    \anonym{\href{https://bugs.webkit.org/show\_bug.cgi?id=188930}{\#188930}}}
    & - & \hi{} & Test262\\
   \bottomrule
  \end{tabular}
\end{table}