### 2020-01-06 msc-thesis/appendix/JAVA/dde.tex
added msc-thesis

% % ----------------------------------------------------------
\chapter{Código Java do DDE2}
\label{sec:dde2java}
% % ----------------------------------------------------------

\begin{lstlisting}[caption={Código em linguagem de programação Java do DDE2},label=dde2java]

/*
 *    DDE.java
 *    Copyright (C) 2015 Maciel, Santos, Barros 
 *    @authors Bruno I. F. Maciel (bifm@cin.ufpe.br)
 *    			Silas Garrido T. de Carvalho Santos (sgtcs@cin.ufpe.br)
 *             	Roberto S. M. Barros (roberto@cin.ufpe.br) 
 *             
 *    @version $Version: 1 $
 *
 *    This program is free software; you can redistribute it and/or modify
 *    it under the terms of the GNU General Public License as published by
 *    the Free Software Foundation; either version 3 of the License, or
 *    (at your option) any later version.
 *
 *    This program is distributed in the hope that it will be useful,
 *    but WITHOUT ANY WARRANTY; without even the implied warranty of
 *    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
 *    GNU General Public License for more details.
 *
 *    You should have received a copy of the GNU General Public License
 *    along with this program. If not, see <http://www.gnu.org/licenses/>.
 */

/**
 * Drift Detection Ensemble (DDE),
 * published as:
 * <p> Bruno I. F. Maciel, Silas Garrido T. C. Santos and Roberto S. M. Barros: 
 *     A Lightweight Concept Drift Detection Ensemble. 
 *     27th IEEE International Conference on Tools with Artificial Intelligence 
 *     (ICTAI) Vietri sul Mare, Italy, November 9-11, 2015</p>
 */

package moa.classifiers.drift;

import moa.classifiers.meta.WEKAClassifier;
import java.util.LinkedList;
import java.util.List;
import moa.classifiers.AbstractClassifier;
import moa.classifiers.Classifier;
import moa.classifiers.core.driftdetection.ChangeDetector;
import moa.core.Measurement;
import moa.options.ClassOption;
import moa.options.IntOption;
import weka.core.Instance;
import weka.core.Utils;

public class DDE2 extends AbstractClassifier  {
    private static final long serialVersionUID = 1L;

    public double[] getVotesForInstance(Instance inst) {
        return this.classifier.getVotesForInstance(inst);
    }

    @Override
    public boolean isRandomizable() {
        return true;
    }

    @Override
    public void getModelDescription(StringBuilder out, int indent) {
        ((AbstractClassifier) this.classifier).getModelDescription(out, indent);
    }

    @Override
    protected Measurement[] getModelMeasurementsImpl() {
        List<Measurement> measurementList = new LinkedList<Measurement>();
        measurementList.add(new Measurement("Change detected", this.changeDetected));
        measurementList.add(new Measurement("Warning detected", this.warningDetected));
        Measurement[] modelMeasurements = ((AbstractClassifier) this.classifier).getModelMeasurements();
        if (modelMeasurements != null) {
            for (Measurement measurement : modelMeasurements) {
                measurementList.add(measurement);
            }
        }
        this.changeDetected = 0;
        this.warningDetected = 0;
        return measurementList.toArray(new Measurement[measurementList.size()]);
    }}

\end{lstlisting}